---
title: "lab2_Whitney_Willoz"
author: "Whitney Willoz"
date: "9/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Reading in the data
```{r}
library(classdata)
data(choco)
str(choco) 
```

1.How are the cocoa percent distributed? Create both a histogram and a boxplot and describe them. Don’t forget to mention outliers, if there are any.

Answer:Based on the histogram and boxplot of the cocoa percent, I was able to find that it is normally distributed with a couple outliers on the high end of percentage. 

```{r}
boxplot(choco$CocoaPercent)
hist(choco$CocoaPercent)
```

2.How does the number of ratings depend on the country of bean origin? Draw a bar chart of the number of ratings by bean origin. [Hint: If your graph is too dense and the labels are not legible, show fewer bean origins.]

Answer:The number of ratings is higher in Peru and Venezuela so higher producing companies probably have more ratings.

```{r}
beanOrigins <- table(choco$CountryBeanOrigin)
beanOrigins1<- sort(beanOrigins, decreasing = TRUE)
beanOrigins2 <- head(beanOrigins1, 10)
barplot(beanOrigins2, las = 2, cex.names = 0.6)
```

3.Do ratings depend on the cacao percentage of a chocolate bar? Create both numerical and graphical summaries, and explain. [Hint: Think about why the scatterplot is not a very good choice for visualization here.]

Answer:The more percent Cocoa was in the chocolate the mean rating for that chocolate went up. There is a positive correlation in the boxplots.

```{r}
summary(choco$CocoaPercent)
summary(factor(choco$CocoaPercent))
summary(factor(choco$Rating))
CocoaPercentonRating <- reorder(factor(choco$CocoaPercent), choco$Rating, FUN=mean)
boxplot(choco$Rating ~ CocoaPercentonRating)

```

## Exploration

4.How do different bean origins compare in terms of the flavor of chocolates? Are beans from some places better than the other? Create both numerical and graphical summaries, and explain.

Answer:Brazil has the creamiest chocolate compared to all other countries and the highest overall rating of the chocolate because of it. Blend has an average rating on the higher side because it has sweet and creamy as a flavor characteristic which are positive characteristics of a chocolate.

```{r}
library(tidyverse)
flavor <- table(choco$Characteristics1)
flavor_filtered <- flavor[flavor > 50]
choco_filtered <- filter(choco, Characteristics1 %in% names(flavor_filtered))

country_filtered <- table(choco$CountryBeanOrigin) %>% 
  sort(decreasing = TRUE) %>%
  head(10) # select the top 10 countries

choco %>% 
  filter(Characteristics1 %in% names(flavor_filtered)) %>% # filter flavor
  filter(CountryBeanOrigin %in% names(country_filtered)) %>% # filter country
  ggplot() +
  geom_bar(aes(x = CountryBeanOrigin, fill = Characteristics1), position = 'fill') + 
  coord_flip() 

tapply(choco$Rating, choco$CountryBeanOrigin, summary) %>%
  head(5)

```

5.brainstorm two analyses you would like to perform. Describe in plain language what comparisons are you making. Create suitable numerical and graphical summaries for your analyses. Describe the results.

Analysis 1:Do chocolates from the US location have a typical lower rating than that of the French?

Answer: On average the French have a higher rating of their chocolate than the US. But only by A very slim bit because there are just more observations of the US that vary.

```{r}
franceL <- choco[choco$Location == "France",]
franceM <- mean(franceL$Rating)
franceM
boxplot(franceL$Rating, xlab = "France", ylab = "Ratings")

USL <- choco[choco$Location == "U.S.A.",]
USM <- mean(USL$Rating)
USM
boxplot(USL$Rating, xlab = "USA", ylab = "Ratings")

```

Analysis 2:Does Amatller (Simon Coll) have a higher rating than Artisan du Chocolat company?

Answer:Amatller (Simon Coll) has a lower average rating and a smaller spread of ratings meaning they must produce lower quality chocolate than Artisan du Chocolat.


```{r}
AmaC <- choco[choco$Company == "Amatller (Simon Coll)", ]
AmaM <- mean(AmaC$Rating)
AmaM
boxplot(AmaC$Rating, xlab = "Amatller (Simon Coll)", ylab = "Ratings")

ArtC <- choco[choco$Company == "Artisan du Chocolat", ]
ArtM <- mean(ArtC$Rating)
ArtM
boxplot(ArtC$Rating, xlab = "Artisan du Chocolat", ylab = "Ratings")

```